---
title: 'Probability and Statistics: Assignment 4'
author: ''
date: '2017-09-05'
slug: probability-and-statistics-assignment-4
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

1.  Modify the code below so that it plots $y=x^3$ over the interval $[-1,3]$.

    ```{r}
    x = seq(-2, 2, by = 0.1)
    y = x^2
    plot(x, y, col = "skyblue", type="l")
    ```

1.  Exercise 4.3

1.  Use simulation to estimate the probabilities you found in the previous problem.

    ```{r}
    # Use this code chunk.
    ```

1.  Consider the random variable that "picks" a number between 0 and 1 with the pdf given by

    \[f(x) = \begin{cases} 6x(1-x), & x\in[0,1] \\ 0 & \text{otherwise}\end{cases}.\]
    
    a.  Plot this density function.
    
    ```{r}
    # Use this code chunk.
    ```
    
    b.  Write code to estimate the integral.
    
    ```{r}
    # Use this code chunk.
    ```
    
    c.  Calculate the exact integral to show that $f(x)$ is, indeed, a probability density function.  
    
1.  Let $X$ be a normal rv with mean 0 and standard deviation 1.

    a.  Plot the density function.
    
    ```{r}
    # Use this code chunk.
    ```
    
    b.  Write code to estimate the integral.
    
    ```{r}
    # Use this code chunk.
    ```
    
    c.  How much of the density is no more than 2 standard deviations away from the mean?
    
    ```{r}
    # Use this code chunk.
    ```
    
    d.  How much of the density is no more than 1 standard deviation away from the mean?
    
    ```{r}
    # Use this code chunk.
    ```
    
    e.  Exercise 4.5.b. Also provide a plot.
    
    ```{r}
    # Use this code chunk.
    ```
