---
title: 'Numerical Methods: Assignment 2'
author: ''
date: '2017-08-26'
slug: numerical-methods-assignment-2
categories: []
tags: []
draft: TRUE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

1.  Let $x=3248$. Use the functions `floor()` and/or `%%` to write an expression that returns the second decimal place in $x$. Make sure your code also works for $x = -3248$.

    ```{r}
    # notice the spacing of this chunk...what is its purpose? 
    x <- 3248
    # write question 1 code inside this chunk
    
    ```

2.  Write an expression that returns the following matrix.  
$\left(\begin{array}{rrr}0 & 2 & 3 \\0 & 5 & 0 \\7 & 0 & 0\end{array}\right)$

    ```{r}
    # write question 2 code inside this chunk
    ```

3.  Produce a vector containing all integers from 1 to 100 that are not divisible by 2, 3, or 7. You want to achieve this by appropriately subsetting the vector `1:100`.

    ```{r}
    # write question 3 code inside this chunk
    ```

4.  Write a simple expression that returns a $10\times 10$ diagonal matrix whose nonzero entries are each equal to 3.

